---
title: "Hands-on Exercise 1"
author: "Leng Yi Xiu"
date: "Aug 25 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

## **Getting Started**

In this hands-on exercise, two R packages will be used:

-   **sf** for importing, managing, and processing geospatial data, and

-   **tidyverse** for performing data science tasks such as importing, wrangling and visualising data.

The code chunk below is to check if sf, tmap and tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## Importing Geospatial Data

In this section, we learn how to import geospatial data into R using *`st_read()`* of sf package.

### Importing polygon feature data in shapefile format

When the input geospatial data is in shapefile format, 2 arguments will be used: *`dsn`* to define the data path and *`layer`* to provide the shapefile name.

```{r}
mpsz = st_read(dsn = "D:/2. SMU - MITB/Term 4/ISSS626 Geospatial/YX-Leng/ISSS626-Geospatial/Hands-on_Ex/Hands-on_Ex01/data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

💡There are a total of 323 multipolygon features and 15 fields in *`mpsz`* simple feature data frame.

💡Bounding box = x extend and y extend of the data.

### Importing polyline feature data in shapefile form

```{r}
cyclingpath = st_read(dsn = "D:/2. SMU - MITB/Term 4/ISSS626 Geospatial/YX-Leng/ISSS626-Geospatial/Hands-on_Ex/Hands-on_Ex01/data/geospatial", 
                         layer = "CyclingPathGazette")
```

💡There are a total of 2,558 features and 2 fields in *`cyclingpath`* linestring feature data frame.

### Importing GIS data in kml format

💡Note that for kml format, the complete file path and the kml file extension is provided.

```{r}
preschool = st_read("D:/2. SMU - MITB/Term 4/ISSS626 Geospatial/YX-Leng/ISSS626-Geospatial/Hands-on_Ex/Hands-on_Ex01/data/geospatial/PreSchoolsLocation.kml")
```

## Checking the Content of a Simple Feature Data Frame

In this section, we learn different ways to retrieve information related to the content of a simple feature data frame.

### Working with *`st_geometry()`*

```{r}
st_geometry(mpsz)
```

### Working with *`glimpse()`*

```{r}
glimpse(mpsz)
```

💡The *`glimpse()`* report reveals the data type of each field.

### Working with *`head()`*

```{r}
head(mpsz, n=5) 
```

## Plotting the Geospatial Data

💡Default plot: multi-plot of all attributes

```{r}
plot(mpsz)
```

💡Plotting only the geometry

```{r}
plot(st_geometry(mpsz))
```

💡Plotting by using specific attribute(s)

```{r}
plot(mpsz["PLN_AREA_N"])
```

## Working with Projection

In this section, we learn how to project a simple feature data frame from one coordinate system to another coordinate system, process also known as **projection transformation**.

### Assigning EPSG code to a simple feature data frame

Common issue that can happen during importing geospatial data into R is the coordinate system of the source data was either

-   missing, or

-   wrongly assigned during the importing process.

```{r}
st_crs(mpsz)
```

::: callout-caution
Code block indicates that EPSG is 9001, which is wrong. Correct EPSG code for svy21 should be **3414**.
:::

💡Assigning the correct EPSG code:

```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

```{r}
st_crs(mpsz3414)
```

### Transforming the projection of preschool from wgs84 to svy21

It is common for us to transform the original data from geographic coordinated system to **projected** coordinated system because the former is not appropriate if the analysis needs to use distance and/or area measurements.

💡 Performing the projection transformation

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

## Importing and Converting an Aspatial Data

In this section, we learn how to import an aspatial data into R environment and save it as a **tibble data frame**, then converting it into a simple feature data frame.

### Importing the data

```{r}
listings <- read_csv("D:/2. SMU - MITB/Term 4/ISSS626 Geospatial/YX-Leng/ISSS626-Geospatial/Hands-on_Ex/Hands-on_Ex01/data/aspatial/listings.csv")
```

```{r}
list(listings) 
```

💡The output reveals a *`listing`* tibble data frame consisting of 3,540 row and 75 columns.

### Creating a simple feature data frame from an aspatial data frame

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

::: {.callout-note appearance="simple"}
-   `coords` argument: provide the column name of the x-coordinates first, then the column name of the y-coordinates

-   `crs` argument: provide the coordinates system in EPSG format\
    4326 = wgs84 Geographic Coordinate System\
    3414 = Singapore SVY21 Projected Coordinate System

-   Search for other country's EPSG code by referring to epsg.io

-   `%>%` used to nest `st_transform()`
:::

```{r}
glimpse(listings_sf)
```

## Geoprocessing with sf package

In this section, we learn how to perform 2 commonly used geoprocessing functions, buffering and point in polygon count.

### 1. Buffering

💡To compute a 5-meter buffer around existing cycling paths, calculate the area of the buffers and `sum()` of Base R to derive total land required.

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
```

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
```

```{r}
sum(buffer_cycling$AREA)
```

### 2. Point in Polygon count

💡Identify pre-schools located inside each Planning Subzone using `st_intersects()`, then calculating number of pre-schools that fall inside each subzone using `length()` of Base R.

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

::: callout-caution
This should not be confused with *`st_intersection()`*
:::

```{r}
summary(mpsz3414$`PreSch Count`)
```

💡 List the planning subzone with the most number of pre-schools

```{r}
top_n(mpsz3414, 1, `PreSch Count`)
```

💡 Calculate the density of pre-schools by planning subzone

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
```

```{r}
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

## Exploratory Data Analysis (EDA)

```{r}
hist(mpsz3414$`PreSch Density`)
```

💡 With customisation - ggplot2

```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

```{r}
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```

# Thematic Mapping and GeoVisualisation with R

Note: **sf**, **tmap** and **tidyverse** packages have been installed in previous sections.

## Importing Data into R

### 1. Geospatial Data

```{r}
mpsz <- st_read(dsn = "D:/2. SMU - MITB/Term 4/ISSS626 Geospatial/YX-Leng/ISSS626-Geospatial/Hands-on_Ex/Hands-on_Ex01/data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz
```

### 2. Attribute Data

```{r}
popdata <- read_csv("D:/2. SMU - MITB/Term 4/ISSS626 Geospatial/YX-Leng/ISSS626-Geospatial/Hands-on_Ex/Hands-on_Ex01/data/aspatial/respopagesextod2011to2020.csv")
```

### 3. Data Preparation

💡 Before a thematic map can be prepared, a data table with Year 2020 values needs to be prepared. This data table should include variables PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY.

-   Data Wrangling

```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[14])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:13])+
rowSums(.[15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

-   Joining the attribute and geospatial data

💡This is an extra step to convert values in PA and SZ to upper case, aligning it to values in SUBZONE_N and PLN_AREA_N.

```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = list(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

💡Join the geographic data and attribute table using planning subzone name.

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

```{r}
write_rds(mpsz_pop2020, "D:/2. SMU - MITB/Term 4/ISSS626 Geospatial/YX-Leng/ISSS626-Geospatial/Hands-on_Ex/Hands-on_Ex01/data/rds/mpszpop2020.rds")
```

## Choropleth Mapping Using *`tmap`*

Choropleth mapping involves the symbolisation of enumeration units, such as countries, provinces, states, counties or census units, using area patterns or graduated colors.

Two approaches can be used to prepare thematic map using **tmap**, they are:

-   Plotting a thematic map quickly by using *`qtm()`*.

-   Plotting highly customisable thematic map by using tmap elements.

### Plotting a choropleth map quickly with *`qtm()`*

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

::: {.callout-caution appearance="simple"}
The disadvantge of *`qtm()`* is that it makes aesthetics of individual layers harder to control.

To draw a high quality cartographic choropleth map as shown in the figure below, *`tmap`*’s drawing elements should be used.
:::

### Creating a choropleth map using tmap's elements

```{r}
tm_shape(mpsz_pop2020)+
tm_fill("DEPENDENCY",
style = "quantile",
palette = "Blues",
title = "Dependency ratio") +
tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
main.title.position = "center",
main.title.size = 1.2,
legend.height = 0.45,
legend.width = 0.35,
frame = TRUE) +
tm_borders(alpha = 0.5) +
tm_compass(type="8star", size = 2) +
tm_scale_bar() +
tm_grid(alpha =0.2) +
tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\\n and Population data from Department of Statistics DOS",
position = c("left", "bottom"))
```

#### Drawing a base map

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

#### Drawing a choropleth map using *`tm_polygons()`*

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

#### Drawing a choropleth map using *`tm_fill()`* and *`tm_border()`*

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  alpha = 1)
```

### Data Classification Methods of **tmap**

**tmap** provides a total ten data classification methods, namely: *`fixed`*, *`sd`*, *`equal`*, *`pretty`* (default), *`quantile`*, *`kmeans`*, *`hclust`*, *`bclust`*, *`fisher`*, and *`jenks`*.

💡 To define a data classification method, the *style* argument of *`tm_fill()`* or *`tm_polygons()`* will be used.

#### Plotting choropleth maps with built-in classification methods

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

::: callout-caution
In the above map, **equal data** classification method was used.

Maps can lie! Prepare choropleth maps using different classification methods supported by *`tmap`* and compare their differences.
:::

#### Plotting choropleth map with custom break

For all the built-in styles, the category breaks are computed internally. In order to override these defaults, the breakpoints can be set explicitly by means of the *`breaks`* argument to the *`tm_fill()`*.

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

💡 With the results above, we set break point at 0.60, 0.70, 0.80 and 0.90. We also need to include minimum and maximum, which is set at 0 and 100. *`breaks`* vector is thus c(0, 0.60, 0.70, 0.80, 0.90, 1.00).

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

### Colour Scheme

#### Using ColourBrewer palette

💡 To change the colour, assign the preferred colour to *`palette`* argument

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```

💡 To reverse the order of the colour shading, add "-" prefix.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```

### Map Layouts

#### Map Legend

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

#### Map Style

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

#### Cartogrphic Furniture

💡 *`tmap`* provides arguments to draw other map furniture (i.e. compass, scale bar and grid lines).

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

💡 To reset back to default style, set it back to "white".

```{r}
tmap_style("white")
```

### **Drawing Small Multiple Choropleth Maps (Facet Maps)**

Facet maps are composed of many maps arrange side-by-side, and sometimes stacked vertically. Small multiple maps enable the visualisation of how spatial relationships change with respect to another variable, such as time.

In **tmap**, these can be plotted in 3 ways:

-   by assigning multiple values to at least one of the asthetic arguments,

-   by defining a group-by variable in *`tm_facets()`*, and

-   by creating multiple stand-alone maps with *`tmap_arrange()`* .

#### By assigning multiple values to at least one of the aesthetic arguments

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

💡 Notice the Blues and Greens

```{r}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

#### By defining a group-by variable in *`tm_facets()`*

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes =TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

#### By creating multiple stand-alone maps with *`tmap_arrange()`*

```{r}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

### Mapping Spatial Object Meeting a Selection Criterion

```{r}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```
