---
title: "In-class Exercise 7"
author: "Leng Yi Xiu"
date: "Oct 14 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

# **Calibrating Hedonic Pricing Model for Private Highrise Property with GWR Method**

In this hands-on exercise, we will learn how to build hedonic pricing models by using GWR methods.

::: {.callout-tip appearance="simple"}
**Geographically weighted regression (GWR)** is a spatial statistical technique that takes non-stationary variables into consideration and models the local relationships between these independent variables and an outcome of interest / dependent variable.
:::

For this exercise, the dependent variable is the resale prices of condominium in 2015. The independent variables are divided into either structural and locational.

## Getting Started

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary, ggstatsplot, sfdep)
```

### Importing Geospatial Data

```{r}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

### Updating CRS information

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

```{r}
st_crs(mpsz_svy21)
```

```{r}
st_bbox(mpsz_svy21) #view extent
```

### Importing aspatial data

```{r}
condo_resale = read_csv("data/aspatial/Condo_resale_2015.csv")
```

```{r}
glimpse(condo_resale)
```

```{r}
head(condo_resale$LONGITUDE) #see the data in XCOORD column
```

```{r}
head(condo_resale$LATITUDE) #see the data in YCOORD column
```

```{r}
summary(condo_resale)
```

### Converting aspatial data frame into a sf object

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%   
  st_transform(crs=3414)
```

```{r}
head(condo_resale.sf)
```

## Exploratory Data Analysis (EDA)

```{r}
ggplot(data=condo_resale.sf, aes(x=`SELLING_PRICE`)) +   geom_histogram(bins=20, color="black", fill="light blue")
```

```{r}
condo_resale.sf <- condo_resale.sf %>%   mutate(`LOG_SELLING_PRICE` = log(SELLING_PRICE))
```

```{r}
ggplot(data=condo_resale.sf, aes(x=`LOG_SELLING_PRICE`)) +   geom_histogram(bins=20, color="black", fill="light blue")
```

💡 Multiple Histogram Plots distribution of variables

```{r}
AREA_SQM <- ggplot(data=condo_resale.sf, aes(x= `AREA_SQM`)) +    
  geom_histogram(bins=20, color="black", fill="light blue")  

AGE <- ggplot(data=condo_resale.sf, aes(x= `AGE`)) +   
  geom_histogram(bins=20, color="black", fill="light blue")  

PROX_CBD <- ggplot(data=condo_resale.sf, aes(x= `PROX_CBD`)) +   
  geom_histogram(bins=20, color="black", fill="light blue")  

PROX_CHILDCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_CHILDCARE`)) +    
  geom_histogram(bins=20, color="black", fill="light blue")  

PROX_ELDERLYCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_ELDERLYCARE`)) +   
  geom_histogram(bins=20, color="black", fill="light blue")  

PROX_URA_GROWTH_AREA <- ggplot(data=condo_resale.sf, aes(x= `PROX_URA_GROWTH_AREA`)) +   
  geom_histogram(bins=20, color="black", fill="light blue")  

PROX_HAWKER_MARKET <- ggplot(data=condo_resale.sf, aes(x= `PROX_HAWKER_MARKET`)) +   
  geom_histogram(bins=20, color="black", fill="light blue")  

PROX_KINDERGARTEN <- ggplot(data=condo_resale.sf, aes(x= `PROX_KINDERGARTEN`)) +   
  geom_histogram(bins=20, color="black", fill="light blue")  

PROX_MRT <- ggplot(data=condo_resale.sf, aes(x= `PROX_MRT`)) +   
  geom_histogram(bins=20, color="black", fill="light blue")  

PROX_PARK <- ggplot(data=condo_resale.sf, aes(x= `PROX_PARK`)) +   
  geom_histogram(bins=20, color="black", fill="light blue")  

PROX_PRIMARY_SCH <- ggplot(data=condo_resale.sf, aes(x= `PROX_PRIMARY_SCH`)) +   
  geom_histogram(bins=20, color="black", fill="light blue")  

PROX_TOP_PRIMARY_SCH <- ggplot(data=condo_resale.sf,                                 
                               aes(x= `PROX_TOP_PRIMARY_SCH`)) +   
  geom_histogram(bins=20, color="black", fill="light blue")  

ggarrange(AREA_SQM, AGE, PROX_CBD, PROX_CHILDCARE, PROX_ELDERLYCARE,            
          PROX_URA_GROWTH_AREA, PROX_HAWKER_MARKET, PROX_KINDERGARTEN, PROX_MRT,           
          PROX_PARK, PROX_PRIMARY_SCH, PROX_TOP_PRIMARY_SCH,             
          ncol = 3, nrow = 4)
```

💡Drawing Statistical Point Map

```{r}
tmap_mode("view")
```

```{r}
tmap_options(check.and.fix = TRUE) 

tm_shape(mpsz_svy21)+   
  tm_polygons() + tm_shape(condo_resale.sf) +     
  tm_dots(col = "SELLING_PRICE",           
          alpha = 0.6,           
          style="quantile") +   
  tm_view(set.zoom.limits = c(11,14))
```

```{r}
tmap_mode("plot")
```

## Hedonic Pricing Modelling in R

### Simple Linear Regression Method

```{r}
condo.slr <- lm(formula=SELLING_PRICE ~ AREA_SQM, data = condo_resale.sf)
```

Dependent variable = *selling price*; independent variable = *AREA_SQM*

```{r}
summary(condo.slr)
```

```{r}
ggplot(data=condo_resale.sf,          
       aes(x=`AREA_SQM`, y=`SELLING_PRICE`)) + 
  geom_point() +   
  geom_smooth(method = lm)
```

### Multiple Linear Regression Method

Code chunk below checks for multicollinearity of independent variables. Quality of regression model will be compromised if highly correlated variables are used.

```{r}
corrplot(cor(condo_resale[, 5:23]), diag = FALSE, order = "AOE",          
         tl.pos = "td", tl.cex = 0.5, method = "number", type = "upper")
```

```{r}
#| fig-width: 12
#| fig-height: 10
ggcorrmat(condo_resale[, 5:23])
```

### Building a hedonic pricing model using multiple linear regression method

```{r}
condo.mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE +                    
                  PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +                   
                  PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET + PROX_KINDERGARTEN +                    
                  PROX_MRT  + PROX_PARK + PROX_PRIMARY_SCH +                   
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_SUPERMARKET +                    
                  PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD +
                  LEASEHOLD_99YR,                     # added back from hands-on Ex07
                data=condo_resale.sf) 

summary(condo.mlr)
```

## Model Assessment: olsrr method

::: callout-note
## Steps

1.  Look at ANOVA results and explain adj. R-Squared results
2.  Describe and elaborate on the variables
3.  Run diagnostic tests - multicollinearity, normality etc.
:::

```{r}
ols_regress(condo.mlr)
```

### Multicollinearity Check

```{r}
ols_vif_tol(condo.mlr)
```

**VIF value is less than 10 - there is no need to exclude the LEASEHOLD_99YR variable.**

### Variable Selection

```{r}
condo_fw_mlr <- ols_step_forward_p(
  condo.mlr,
  p_val = 0.05,
  details = TRUE)
```

```{r}
#| fig-width: 12
#| fig-height: 10

plot(condo_fw_mlr)
```

### Test for Non-Linearity

```{r}
ols_plot_resid_fit(condo_fw_mlr$model)
```

### Test for Normality

```{r}
ols_plot_resid_hist(condo_fw_mlr$model)
```

```{r}
ols_test_normality(condo_fw_mlr$model)
```

### Testing for Spatial Autocorrelation

```{r}
mlr_output <- as.data.frame(condo_fw_mlr$model$residuals) %>%
  rename('FW_MLR_RES' = 'condo_fw_mlr$model$residuals')
```

Next, we will join the newly created data frame with *condo_resale_sf* object.

```{r}
condo_resale.sf <- cbind(condo_resale.sf,             # Use cbind because there is no unique identifier in the tables
                         mlr_output$FW_MLR_RES) %>%
  rename('MLR_RES_2' = 'mlr_output.FW_MLR_RES')
```

```{r}
tmap_mode("view")
tm_shape(mpsz)+
  tmap_options(check.and.fix = TRUE) +
  tm_polygons(alpha = 0.4) +
tm_shape(condo_resale.sf) +
  tm_dots(col = "MLR_RES_2",
          alpha = 0.6,
          style = "quantile")
tmap_mode("plot")
```

### Moran's I Test

```{r}
condo_resale.sf <- condo_resale.sf %>%
  mutate(nb = st_knn(geometry, k=6,
                     longlat = FALSE),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1)
```

```{r}
global_moran_perm(condo_resale.sf$MLR_RES_2,
                  condo_resale.sf$nb,
                  condo_resale.sf$wt,
                  alternative = "two.sided",
                  nsim = 99)
```

------------------------------------------------------------------------
## Building Hedonic Pricing Models using GWmodel

### Building Fixed Bandwidth GWR Model

💡 Computing fixed bandwidth

```{r}
bw.fixed <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE + PROX_CBD + 
                     PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + 
                     PROX_MRT   + PROX_PARK + PROX_PRIMARY_SCH + 
                     PROX_SHOPPING_MALL + PROX_BUS_STOP + NO_Of_UNITS + 
                     FAMILY_FRIENDLY + FREEHOLD, 
                   data=condo_resale.sf, 
                   approach="CV", 
                   kernel="gaussian", 
                   adaptive=FALSE, 
                   longlat=FALSE)
```

💡Calibrating the GWR model using fixed bandwidth and gaussian kernel

```{r}
gwr.fixed <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE + PROX_CBD + 
                         PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + 
                         PROX_MRT   + PROX_PARK + PROX_PRIMARY_SCH + 
                         PROX_SHOPPING_MALL + PROX_BUS_STOP + NO_Of_UNITS + 
                         FAMILY_FRIENDLY + FREEHOLD, 
                       data=condo_resale.sf, 
                       bw=bw.fixed, 
                       kernel = 'gaussian', 
                       longlat = FALSE)
```

```{r}
gwr.fixed
```

### Building Adaptive Bandwidth GWR Model

💡 Computing adaptive bandwidth

```{r}
bw.adaptive <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
                        PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    + 
                        PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                        PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + PROX_BUS_STOP + 
                        NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                      data=condo_resale.sf, 
                      approach="CV", 
                      kernel="gaussian", 
                      adaptive=TRUE, 
                      longlat=FALSE)
```

💡 Constructing the adaptive bandwidth gwr model

```{r}
gwr.adaptive <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                            PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE + 
                            PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                            PROX_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_BUS_STOP + 
                            NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                          data=condo_resale.sf, bw=bw.adaptive, 
                          kernel = 'gaussian', 
                          adaptive=TRUE, 
                          longlat = FALSE)
```

```{r}
gwr.adaptive
```

### Visualising GWR Output

Condition Number: Results associated with condition numbers larger than 30, may be unreliable.

Local R2: Values range between 0.0 and 1.0. Very low values indicate the local model is performing poorly.

Residuals: The fitted y values are subtracted from the observed y values.

Coefficient Standard Error: Confidence in those estimates are higher when standard errors are small in relation to the actual coefficient values. Large standard errors may indicate problems with local collinearity.


### Converting SDF into sf data.frame

```{r}
condo_resale.sf.adaptive <- st_as_sf(gwr.adaptive$SDF) %>%
  st_transform(crs=3414)
```

```{r}
condo_resale.sf.adaptive.svy21 <- st_transform(condo_resale.sf.adaptive, 3414)
condo_resale.sf.adaptive.svy21
```

```{r}
gwr.adaptive.output <- as.data.frame(gwr.adaptive$SDF)
condo_resale.sf.adaptive <- cbind(condo_resale.sf, as.matrix(gwr.adaptive.output))
```

```{r}
glimpse(condo_resale.sf.adaptive)
```

```{r}
summary(gwr.adaptive$SDF$yhat)
```
