---
title: "Hands-on Exercise 8"
author: "Leng Yi Xiu"
date: "Oct 20 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

# **Geographically Weighted Predictive Models**

In this in-class exercise, we will learn how to build predictive model by using geographical random forest method. By the end of this hands-on exercise, we will acquire the skills of:

-   preparing training and test data sets by using appropriate data sampling methods,

-   calibrating predictive models by using both geospatial statistical learning and machine learning methods,

-   comparing and selecting the best model for predicting the future outcome,

-   predicting the future outcomes by using the best model calibrated.

### Installing and Loading R packages

```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse)
```

### Reading data file to rds

```{r}
mdata <- read_rds("data/model/mdata.rds")
```

### Data Sampling

```{r}
set.seed(1234)
resale_split <- initial_split(mdata, 
                              prop = 0.65/10,)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

```{r}
write_rds(train_data, "data/model/train_data.rds")
write_rds(test_data, "data/model/test_data.rds")
```

### Computing Correlation Matrix

```{r}
mdata_nogeo <- mdata %>%
  st_drop_geometry()
corrplot::corrplot(cor(mdata_nogeo[, 2:17]), 
                   diag = FALSE, 
                   order = "AOE",
                   tl.pos = "td", 
                   tl.cex = 0.5, 
                   method = "number", 
                   type = "upper")
```

::: callout-important
## Observations

All the correlation values are \< 0.8, hence there is no signs of multicollinearity.
:::

## Building a non-spatial multiple linear regression

```{r}
train_data <- read_rds("data/model/train_data.rds")
test_data <- read_rds("data/model/test_data.rds")
```

```{r}
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
summary(price_mlr)
```

```{r}
write_rds(price_mlr, "data/model/price_mlr.rds" ) 
```

## GWR Predictive Method

In this section, we will learn how to calibrate a model to predict HDB resale price with geographically weighted regression method of GWmodel package.

### Converting the sf data frame to spatial point data frame

```{r}
train_data_sp <- as_Spatial(train_data)
train_data_sp
```

### Computing adaptive bandwidth

```{r}
bw_adaptive <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=train_data_sp,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```

```{r}
write_rds(bw_adaptive, "data/model/bw_adaptive.rds")
```

### Constructing the adaptive bandwidth gwr model

```{r}
bw_adaptive <- read_rds("data/model/bw_adaptive.rds")
```

```{r}
gwr_adaptive <- gwr.basic(formula = resale_price ~
                            floor_area_sqm + storey_order +
                            remaining_lease_mths + PROX_CBD + 
                            PROX_ELDERLYCARE + PROX_HAWKER +
                            PROX_MRT + PROX_PARK + PROX_MALL + 
                            PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                            WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                            WITHIN_1KM_PRISCH,
                          data=train_data_sp,
                          bw=bw_adaptive, 
                          kernel = 'gaussian', 
                          adaptive=TRUE,
                          longlat = FALSE)
```

```{r}
write_rds(gwr_adaptive, "data/model/gwr_adaptive.rds")
```

### Retrive gwr output object

```{r}
gwr_adaptive <- read_rds("data/model/gwr_adaptive.rds")
```

```{r}
gwr_adaptive
```

### Converting the test data from sf data frame to Spatial point data frame

```{r}
test_data_sp <- test_data %>%
  as_Spatial()
test_data_sp
```


## Preparing coordinates data

### Extracting coordinates data

```{r}
coords <- st_coordinates(mdata)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```

```{r}
coords_train <- write_rds(coords_train, "data/model/coords_train.rds" )
coords_test <- write_rds(coords_test, "data/model/coords_test.rds" )
```

### Dropping geometry field

```{r}
train_data <- train_data %>% 
  st_drop_geometry()
```

### Calibrating Random Forest Model

```{r}
set.seed(1234)
rf <- ranger(resale_price ~ floor_area_sqm + storey_order + 
               remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE + 
               PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL + 
               PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
               WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
               WITHIN_1KM_PRISCH,
             data=train_data)
rf
```

```{r}
write_rds(rf, "data/model/rf.rds")
```

```{r}
rf <- read_rds("data/model/rf.rds")
rf
```

## Calibrating Geographical Random Forest Model

### Calibrating using training data

```{r}
set.seed(1234)
gwRF_adaptive <- grf(formula = resale_price ~ floor_area_sqm + storey_order +
                       remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE +
                       PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL +
                       PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                       WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                       WITHIN_1KM_PRISCH,
                     dframe=train_data, 
                     bw=55,
                     kernel="adaptive",
                     coords=coords_train)
```

```{r}
write_rds(gwRF_adaptive, "data/model/gwRF_adaptive.rds")
```

```{r}
gwRF_adaptive <- read_rds("data/model/gwRF_adaptive.rds")
```

### Predicting using test data

```{r}
test_data <- cbind(test_data, coords_test) %>%
  st_drop_geometry()
```

💡 **Predicting with test data**

```{r}
gwRF_pred <- predict.grf(gwRF_adaptive, 
                           test_data, 
                           x.var.name="X",
                           y.var.name="Y", 
                           local.w=1,
                           global.w=0)
```

```{r}
GRF_pred <- write_rds(gwRF_pred, "data/model/GRF_pred.rds")
```

💡 **Convert the output into a data frame**

```{r}
GRF_pred <- read_rds("data/model/GRF_pred.rds")
GRF_pred_df <- as.data.frame(GRF_pred)
```

```{r}
test_data_p <- cbind(test_data, GRF_pred_df)
```

```{r}
write_rds(test_data_p, "data/model/test_data_p.rds")
```

### Calculating Root Mean Square Error

```{r}
rmse(test_data_p$resale_price, 
     test_data_p$GRF_pred)
```

### Visualising the predicted values

```{r}
ggplot(data = test_data_p,
       aes(x = GRF_pred,
           y = resale_price)) +
  geom_point()
```
