---
title: "In-class Exercise 2"
author: "Leng Yi Xiu"
date: "Sep 02 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

## Installing *`maptools`*

```{r}
#| eval: false
install.packages("maptools", 
                 repos = "https://packagemanager.posit.co/cran/2023-10-13")
```

```{r}
pacman::p_load(sf, raster, spatstat, tmap, tidyverse)
```

```{r}
childcare_sf <- st_read("D:/2. SMU - MITB/Term 4/ISSS626 Geospatial/YX-Leng/ISSS626-Geospatial/Hands-on_Ex/Hands-on_Ex02/data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
sg_sf <- st_read(dsn = "D:/2. SMU - MITB/Term 4/ISSS626 Geospatial/YX-Leng/ISSS626-Geospatial/Hands-on_Ex/Hands-on_Ex02/data", layer="CostalOutline")
```

```{r}
mpsz_sf <- st_read(dsn = "D:/2. SMU - MITB/Term 4/ISSS626 Geospatial/YX-Leng/ISSS626-Geospatial/Hands-on_Ex/Hands-on_Ex02/data", 
                layer = "MP14_SUBZONE_WEB_PL")
```

## How to use st_union()

```{r}
mpsz_sf_simplified <- st_simplify(mpsz_sf, dTolerance = 0.01)
sg_sf <- st_union(mpsz_sf_simplified)

```

```{r}
plot(sg_sf, col = "white", main = "Union of sg_sf and mpsz_sf")
```


## Monte Carlo Simulation

::: {.callout-important appearance="simple"}
To ensure reproducibility, important to set seed before using spatial functions to involve monte carlo simulation.
:::

```{r}
set.seed(1234)
```

## Edge Correction

If we are looking at data points near the boundaries of a study region, edge correction is essential to eliminate bias and ensure accurate estimates.

i.e. If we are looking at the entire Singapore map, then edge correction not needed. However, if we suspect edge effects, then should test for edge correction.
