---
title: "Take-home Exercise 3: Modelling Geography of Financial Inclusion with Geographically Weighted Methods"
author: "Leng Yi Xiu"
date: "Nov 2 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

# 1 Overview

Tanzania is one of the Africa’s fastest growing economies with nearly 7 percent annual national GDP growth since 2000. Despite the robust economic growth, widespread poverty persists. **Approximately 49% of Tanzania’s population lives below the international extreme poverty line.** This could be due to the lack of financial inclusion - the availability and equality of opportunities to access financial services such as banking, loan, equity and insurance products.

## 1.1 Objective

The objective of this exercise is to build an exploratory model that determines factors affecting financial inclusion by using **non-spatial regression** and **geographically weighted regression methods**.

Examples of **independent** factors include demographics (rural vs urban), marital status, gender, financial education, access to technology etc.

Examples of **dependent** factors include savings, financial situation, free access to banks, insurance coverage etc.

------------------------------------------------------------------------

# 2 Getting Started

The code chunk below installs and loads **olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary** packages into R environment:

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary)
```

------------------------------------------------------------------------

# 3 Data Preparation

For the purpose of this exercise, we will be using two data sets:

1.  District level boundary GIS data for Tanzania

2.  FinScope Tanzania 2023 Survey results

## 3.1 Importing data into R Environment

::: panel-tabset
## Geospatial

```{r}
GB = st_read(dsn = "data/rawdata/geospatial", layer = "geoBoundaries-TZA-ADM2")
```

💡 Updating CRS information

```{r}
st_crs(GB) <- 4210
GB <- st_transform(GB, crs = 32736)
head(GB, 5)
```

## Aspatial

```{r}
Tanz_raw = read_csv("data/rawdata/aspatial/FinScope Tanzania 2023_Individual Main Data_FINAL.csv")
```

```{r}
head(Tanz_raw, 5)
```

💡 Count number of unique dist_name

```{r}
Tanz_raw %>%
  count(dist_name) %>%
  print()
```
:::

## 3.2 Geospatial: Converting Multipolygon to Polygon

As the district level boundary GIS data in multipolygon form will cause overlays or intersections, it is more intuitive to calculate the area of each polygon and group them together by the unique district name and then select the largest polygon by area.

💡 Calculate the area of each polygon

```{r}
sf_GB <- GB %>%
  st_cast("POLYGON") %>%
  mutate(area = st_area(.))
```

💡 Group and select the largest polygon by area

```{r}
GB_clean <- sf_GB %>%
  group_by(shapeName) %>%
  filter(area == max(area)) %>%
  ungroup() %>%
  select(-area) %>%
  select(shapeName)
```

## 3.3 Aspatial: Filtering columns for information required

As the FinScope Tanzania 2023 dataset contains 721 variables and not all variables will be used in our analysis, we will filter for the necessary independent and dependent variables based to analyse on:

\(a\) how cluster type, gender and access to technology affects emergency savings, and

\(b\) how cluster type, gender and access to technology affects borrowing.

```{r}
Tanz_filter <- Tanz_raw %>%
  select(dist_name, clustertype, c8c, c9, c23__1,c23__2,,e_7_n_3,F4_1__3,F4_1__5,F4_1__8,F4_1__13,F4_1__14,
         g_2_3,g_5_1__1,g_5_1__2,g_5_1__3,g_5_1__5,g_5_1__6,g_5_1__7,
         g_5_1__8,g_5_1__9,g_5_1__10,g_5_1__11,g_5_1__12,g_5_1__13,
         ins3_1__3,ins3_1__9,Household_weight,population_wt)
```

## 3.4 Aspatial: Rename Columns

```{r}
Tanz_filter <- Tanz_filter %>%
  rename(
    Age = c8c,
    Gender = c9, 
    Access_MobilePhone = c23__1,
    Access_Internet = c23__2,
    Savings_Emerg = e_7_n_3,
    Savings_SACCOS = F4_1__3,
    Savings_MobileWallet = F4_1__5,
    Savings_Group = F4_1__8,
    Savings_Cash = F4_1__13,
    Savings_Lifestock = F4_1__14,
    Borrow_Amount = g_2_3,
    Borrow_Bank = g_5_1__1,
    Borrow_MFI = g_5_1__2,
    Borrow_SACCOS = g_5_1__3,
    Borrow_Gov = g_5_1__5,
    Borrow_MM = g_5_1__6,
    Borrow_Pension = g_5_1__7,
    Borrow_Employer = g_5_1__8,
    Borrow_Fam_Pay = g_5_1__9,
    Borrow_Fam_NoPay = g_5_1__10,
    Borrow_SavGroup = g_5_1__11,
    Borrow_ML = g_5_1__12,
    Borrow_Religious = g_5_1__13,
    Insurance_Health = ins3_1__3,
    Insurance_Life = ins3_1__9
    )
```

## 3.5 Aspatial: Convert Yes/No to a numeric vector

```{r}
Tanz_filter$clustertype <- as.numeric(Tanz_filter$clustertype == "Urban")
Tanz_filter$Gender <- as.numeric(Tanz_filter$Gender == "Female")
Tanz_filter$Access_MobilePhone <- as.numeric(Tanz_filter$Access_MobilePhone == "Yes")
Tanz_filter$Access_Internet <- as.numeric(Tanz_filter$Access_Internet == "Yes")
Tanz_filter$Savings_Emerg <- as.numeric(Tanz_filter$Savings_Emerg == "Yes")
Tanz_filter$Savings_SACCOS <- as.numeric(Tanz_filter$Savings_SACCOS == "Yes")
Tanz_filter$Savings_MobileWallet <- as.numeric(Tanz_filter$Savings_MobileWallet == "Yes")
Tanz_filter$Savings_Group <- as.numeric(Tanz_filter$Savings_Group == "Yes")
Tanz_filter$Savings_Cash <- as.numeric(Tanz_filter$Savings_Cash == "Yes")
Tanz_filter$Savings_Lifestock <- as.numeric(Tanz_filter$Savings_Lifestock == "Yes")
Tanz_filter$Borrow_Bank <- as.numeric(Tanz_filter$Borrow_Bank == "Yes")
Tanz_filter$Borrow_MFI <- as.numeric(Tanz_filter$Borrow_MFI == "Yes")
Tanz_filter$Borrow_SACCOS <- as.numeric(Tanz_filter$Borrow_SACCOS == "Yes")
Tanz_filter$Borrow_Gov <- as.numeric(Tanz_filter$Borrow_Gov == "Yes")
Tanz_filter$Borrow_MM <- as.numeric(Tanz_filter$Borrow_MM == "Yes")
Tanz_filter$Borrow_Pension <- as.numeric(Tanz_filter$Borrow_Pension == "Yes")
Tanz_filter$Borrow_Employer <- as.numeric(Tanz_filter$Borrow_Employer == "Yes")
Tanz_filter$Borrow_Fam_Pay <- as.numeric(Tanz_filter$Borrow_Fam_Pay == "Yes")
Tanz_filter$Borrow_Fam_NoPay <- as.numeric(Tanz_filter$Borrow_Fam_NoPay == "Yes")
Tanz_filter$Borrow_SavGroup <- as.numeric(Tanz_filter$Borrow_SavGroup == "Yes")
Tanz_filter$Borrow_ML <- as.numeric(Tanz_filter$Borrow_ML == "Yes")
Tanz_filter$Borrow_Religious <- as.numeric(Tanz_filter$Borrow_Religious == "Yes")
Tanz_filter$Insurance_Health <- as.numeric(Tanz_filter$Insurance_Health == "Yes")
Tanz_filter$Insurance_Life <- as.numeric(Tanz_filter$Insurance_Life == "Yes")
```

```{r}
Tanz_filter$Borrow_Amount <- ifelse(is.na(Tanz_filter$Borrow_Amount), 0, Tanz_filter$Borrow_Amount)
Tanz_filter$Borrow_Amount <- as.numeric(Tanz_filter$Borrow_Amount)
```

```{r}
Tanz_filter[is.na(Tanz_filter)] <- 0
```

```{r}
head(Tanz_filter,5)
```

## 3.6 Aspatial: Replace dist_name to avoid empty polygons

::: callout-important
## Assumptions

For the dist_name that cannot match to the GIS data for Tanzania, it will be replaced by the values of its Mother District (for coast) or the nearest city.
:::

```{r}
Tanz_join <- Tanz_filter %>%
  mutate(dist_name = case_when(
    dist_name == "Tanganyika" ~ "Tanga Urban",
    dist_name == "Kigamboni" ~ "Temeke",
    dist_name == "Arumeru" & clustertype == "1" ~ "Arusha Urban",
    dist_name == "Arumeru" & clustertype == "0" ~ "Arusha",
    dist_name == "Butiama" ~ "Butiam",
    dist_name == "Dodoma" ~ "Dodoma Urban",
    dist_name == "Tanga" ~ "Tanga Urban",
    dist_name == "Malinyi" ~ "Morogoro",
    dist_name == "Magharibi B"~ "Magharibi",
    dist_name == "Magharibi A"~ "Magharibi",
    dist_name == "Ubungo"~ "Kinondoni",
    dist_name == "Tabora"~ "Tabora Urban",
    dist_name == "Kibiti"~ "Mkuranga",
    TRUE ~ dist_name  # Keep all other values unchanged
  ))
```

## 3.7 Perform relational join

```{r}
Tanz_final <- left_join(Tanz_join,GB_clean,
                  by = c("dist_name" = "shapeName"), relationship = "many-to-many")
```

💡 Check for any empty polygons and class

```{r}
empty_polygons <- Tanz_final %>%
  filter(st_is_empty(geometry)) %>%
  select(dist_name) %>%
  distinct()
```

```{r}
class(Tanz_final)
```

## 3.8 Writing, reading data file to rds

```{r}
write_rds(Tanz_final, "data/rds/Tanz_final.rds")
```

```{r}
Tanz_final <- read_rds("data/rds/Tanz_final.rds")
```

------------------------------------------------------------------------

# 4 Non-Spatial Regression Methods

## 4.1 Check for multi-collinearity

```{r}
Tanz_corr <- Tanz_final %>%
  st_drop_geometry()
corrplot::corrplot(cor(Tanz_corr[,2:27]), 
                   diag = FALSE, 
                   order = "AOE",
                   tl.pos = "td", 
                   tl.cex = 0.5, 
                   method = "number", 
                   type = "upper")
```

::: callout-important
## Observations

Since all the correlation values are **\< 0.8**, there is no signs of multi-collinearity.
:::

## 4.2 Build a non-spatial multi-linear regression

### 4.2.1 How is borrowing amount affected by various independent variables?

```{r}
## need to log the borrowing amount to normalize extreme values
Tanz_final$Borrow_Amount <- log(Tanz_final$Borrow_Amount + 1)
```

```{r}
Borrowing_mlr <- lm(
  Borrow_Amount ~ clustertype + Age + Gender + Access_MobilePhone + Access_Internet + Borrow_Bank + Borrow_MFI + 
  Borrow_SACCOS + Borrow_Gov + Borrow_MM + Borrow_Pension + Borrow_Employer + Borrow_Fam_Pay + Borrow_Fam_NoPay +  
  Borrow_SavGroup + Borrow_ML + Borrow_Religious,
  data=Tanz_final)

summary(Borrowing_mlr)
```

::: callout-important
## Observations

Adjusted R-squared of **\~80% (0.7963)** indicates a strong model fit and no overfitting.

The model shows that the source of borrowing strongly influenced the borrowing amounts. Tanzanians tend to borrow from micro-finance institutions (MFI) and family/religious groups, but less from Pension (negative estimate). The negative estimate for gender also indicates that males borrow less than females (one-hot encoding female = 1).

The significance of access to mobile phones to borrowing amount suggests that financial accessibility may be linked to mobile connectivity.
:::

::: panel-tabset
## Multicollinearity Check

```{r}
ols_vif_tol(Borrowing_mlr)
```

## Non-Linearity Test

```{r}
ols_plot_resid_fit(Borrowing_mlr)
```

## Normality Assumption

```{r}
ols_plot_resid_hist(Borrowing_mlr)
```

## Spatial Autocorrelation Test

```{r}
mlr.output <- as.data.frame(Borrowing_mlr$residuals) 
```

```{r}
Borrowing.res.sf <- cbind(Tanz_final, 
                          MLR_RES = Borrowing_mlr$residuals)
```

```{r}
Borrowing.sf <- st_as_sf(Borrowing.res.sf)
Borrowing.sp <- as_Spatial(Borrowing.sf)
Borrowing.sp
```
:::

```{r}
tmap_mode("view")
tm_shape(GB_clean)+
  tmap_options(check.and.fix = TRUE) +
  tm_polygons(alpha = 0.4) +
tm_shape(Borrowing.sp) +
  tm_dots(col = "MLR_RES",
          alpha = 0.6,
          style = "quantile")
tmap_mode("plot")
```

### 4.2.2 How is emergency savings affected by various independent variables?

```{r}
Savings_mlr <- lm(
  Savings_Emerg ~ clustertype + Age + Gender + Access_MobilePhone + Access_Internet + Savings_SACCOS + 
  Savings_MobileWallet + Savings_Group + Savings_Cash + Savings_Lifestock + Insurance_Health + Insurance_Life,
  data=Tanz_final)

summary(Savings_mlr)
```

::: callout-important
## Observations

Very low p-value (\< 2.2e-16) indicating that the model is statistically significant.

-   **Gender**: The negative estimate (-0.04, p \< 0.001) suggests that males tend to save less for emergencies.

-   **Age**: There’s a slight but significant negative effect of age on emergency savings, possibly indicating a reduced focus on emergency savings as age increases.

-   **Insurance (Health and Life)**: high p-values indicative that there is no significant relationship with emergency savings.

The low R-squared value suggests that emergency savings might be influenced by factors not captured in this model, such as level of inome, financial literacy, or household characteristics. Further exploration with additional variables could provide more insights into determinants of emergency savings.
:::

::: panel-tabset
## Multicollinearity Check

```{r}
ols_vif_tol(Savings_mlr)
```

## Non-Linearity Test

```{r}
ols_plot_resid_fit(Savings_mlr)
```

## Normality Assumption

```{r}
ols_plot_resid_hist(Savings_mlr)
```

## Spatial Autocorrelation Test

```{r}
mlr.output <- as.data.frame(Savings_mlr$residuals) 
```

```{r}
Savings.res.sf <- cbind(Tanz_final, 
                          MLR_RES = Savings_mlr$residuals)
```

```{r}
Savings.sf <- st_as_sf(Savings.res.sf)
Savings.sp <- as_Spatial(Savings.sf)
Savings.sp
```
:::

```{r}
tmap_mode("view")
tm_shape(GB_clean)+
  tmap_options(check.and.fix = TRUE) +
  tm_polygons(alpha = 0.4) +
tm_shape(Savings.sp) +
  tm_dots(col = "MLR_RES",
          alpha = 0.6,
          style = "quantile")
tmap_mode("plot")
```

# 5 Geographically Weighted Regression Methods

**Geographically weighted regression (GWR)** is a spatial statistical technique that takes non-stationary variables into consideration and models the local relationships between these independent variables and an outcome of interest / dependent variable.

## 5.1 Building Adaptive Bandwidth GWR Model for Borrowing

```{r}
bw.adaptive <- bw.gwr(formula = Borrow_Amount ~ clustertype + Age + Gender + Access_MobilePhone + Access_Internet + Borrow_Bank + Borrow_MFI + 
                     Borrow_SACCOS + Borrow_Gov + Borrow_MM + Borrow_Pension + Borrow_Employer + Borrow_Fam_Pay + Borrow_Fam_NoPay +  
                     Borrow_SavGroup + Borrow_ML + Borrow_Religious, 
                     data=Borrowing.sp, 
                     approach="CV", 
                     kernel="gaussian", 
                     adaptive=TRUE, 
                     longlat=FALSE)
```

💡Calibrating the GWR model using adaptive bandwidth and gaussian kernel

```{r}
gwr.adaptive <- gwr.basic(formula = Borrow_Amount ~ clustertype + Age + Gender + Access_MobilePhone + Access_Internet + Borrow_Bank + Borrow_MFI + 
                      Borrow_SACCOS + Borrow_Gov + Borrow_MM + Borrow_Pension + Borrow_Employer + Borrow_Fam_Pay + Borrow_Fam_NoPay +  
                      Borrow_SavGroup + Borrow_ML + Borrow_Religious, 
                      data=Borrowing.sp,
                      bw=bw.adaptive, 
                      kernel = 'gaussian', 
                      adaptive=TRUE, 
                      longlat = FALSE)
```

```{r}
gwr.adaptive
```

::: callout-important
## Observations

**Coefficient Estimates:**

-   **Clustertype**: Ranges from -0.149 to 0.406, showing a negative influence in some areas and positive in others. This suggests that certain clusters might either restrict or encourage borrowing differently based on location.

-   **Gender**: Consistently negative - indicative that across locations, males may have lower borrowing amounts than females, with the effect stronger in some areas.

-   **Access to Mobile and Internet**: Generally positive effects, suggesting that access to technology generally increases borrowing, although there are areas where internet access has a slight negative effect.

**Borrowing Sources with Divergent Effects:**

-   **Borrow_Gov**: Ranges widely from -5.30 to 12.53, showing that in some areas, government borrowing is linked to lower borrowing amounts, while in others, it has a high positive effect.

-   **Borrow_Pension**: Large variability (-43.57 to 9.69), suggesting that pension borrowing strongly decreases borrowing amounts in some areas while having a positive effect in others.

These results suggest that interventions aimed at influencing borrowing behavior could benefit from being **spatially targeted**. For example, policies could focus on enhancing mobile and internet access in regions where they are most effective or promoting specific borrowing sources depending on local impact.
:::

## 5.2 Building Adaptive Bandwidth GWR Model for Emergency Savings

```{r}
bw.adaptive2 <- bw.gwr(formula = Savings_Emerg ~ clustertype + Age + Gender + Access_MobilePhone + Access_Internet + Savings_SACCOS + 
                       Savings_MobileWallet + Savings_Group + Savings_Cash + Savings_Lifestock + Insurance_Health + Insurance_Life, 
                       data=Savings.sp, 
                       approach="CV", 
                       kernel="gaussian", 
                       adaptive=TRUE, 
                       longlat=FALSE)
```

```{r}
gwr.adaptive2 <- gwr.basic(formula = Savings_Emerg ~ clustertype + Age + Gender + Access_MobilePhone + Access_Internet + Savings_SACCOS + 
                       Savings_MobileWallet + Savings_Group + Savings_Cash + Savings_Lifestock + Insurance_Health + Insurance_Life, 
                       data=Savings.sp,
                       bw=bw.adaptive, 
                       kernel = 'gaussian', 
                       adaptive=TRUE, 
                       longlat = FALSE)
```

```{r}
gwr.adaptive2
```
